---
title: "Assignment 1 sok1005"
format: html
editor: visual
---

## Assignment

Download the four data series from: **https://www.drroyspencer.com/latest-global-temperatures/**
Use only the Global temperature from each dataset (i.e., the third data column).

calculate the 12-month(right-aligned) moving average
calculate a new variable called **Average**, which is the average of the four 12-month moving averages.
Make a plot of the four-atmosphere locations and their **Average** in the same plot.

use data from January 1980 and onwards. 

## Code

Setup

```{r}
rm(list = ls())
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(zoo))


# A custome theme I use

hod_theme_colours <- c("#ffa822","#1ac0c6","#ff6150","#30B66A","#B06AFF","#FF21E1")

hod_theme <- function(alter_x_axis_colour = FALSE,
                      ledgend_pos = "right") {
  
  colour_text <- "#b8aba7"
  colour_text_light <- "#FFE9E3"
  colour_box <- "#2b2b2b"
  colour_lines <- "#827974"
  colour_main <- "#E9D5DA"
  colour_background <- "#2b2b2b"
  colour_palette_standard <- c("#ffa822", "#ff6150","#1ac0c6")
  colour_text_x_axis <- colour_text
  
  if(alter_x_axis_colour){
     colour_text_x_axis <- c(colour_text,colour_text_light)
  }
    
  theme(
    # Border and margin
    plot.margin = margin(0.4, 0.6, 0.4, 0.4, "cm"),
    panel.border = element_rect(colour = colour_box, fill = NA, linetype = 2),
    
    # background
    panel.background = element_rect(fill = colour_background),
    plot.background = element_rect(fill = colour_background),
    strip.background = element_rect(fill = colour_background),
    
    rect = element_rect(fill = colour_background, colour = "white",
                        size = 0.5, linetype = 1),
    
    # grid
    panel.grid.major.x = element_line(colour = colour_lines, linetype = 5, size = 0.25),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y =  element_line(colour = colour_lines, linetype = 5, size = 0.25),
    panel.grid.minor.y = element_blank(),
    axis.ticks = element_line(colour = colour_text),
    
    # line
    line = element_line(colour = colour_palette_standard),
    
    # text
    text = element_text(colour = colour_text, face = "italic", family = "mono"),
    plot.title = element_text(colour = colour_text_light, face = "bold", family = "mono"),
    plot.subtitle = element_text(colour = colour_text, family = "mono", hjust = 0.01),
    plot.caption = element_text(colour = colour_text, face = "italic", family = "mono"),
    axis.text.y = element_text(colour = colour_text, face = "italic", family = "mono"),
    axis.text.x = element_text(colour = colour_text_x_axis, face = "italic", family = "mono"),
    axis.title = element_text(colour = colour_text_light, family = "mono"),
    plot.tag = element_text(colour = colour_text, face = "italic", family = "mono"),
    legend.text = element_text(color = colour_text,face = "italic", family = "mono"),
    legend.title = element_text(colour = colour_text, face = "italic", family = "mono"),
    strip.text = element_text(colour = colour_text_light, face = "bold", family = "mono"),
    
    # legend
    legend.position = ledgend_pos,
    legend.key = element_rect(fill = colour_background)
  )
}

```

Importing of data

```{r echo = FALSE, message=FALSE, warning=FALSE}
df_Lower_Troposphere_raw <- read_table("https://www.nsstc.uah.edu/data/msu/v6.0/tlt/uahncdc_lt_6.0.txt")
df_Mid_Troposphere_raw <- read_table("https://www.nsstc.uah.edu/data/msu/v6.0/tmt/uahncdc_mt_6.0.txt")
df_Tropopause_raw <- read_table("https://www.nsstc.uah.edu/data/msu/v6.0/ttp/uahncdc_tp_6.0.txt")
df_Lower_Stratosphere_raw <- read_table("https://www.nsstc.uah.edu/data/msu/v6.0/tls/uahncdc_ls_6.0.txt")
```

Clean Data

```{r}
clean_data <- function(df,mo_range = c(12:12),year_range = c(1980:2023),new_globe_name = "Globe"){
  return(df %>% 
    filter(Year %in% year_range,
           Mo %in% mo_range) %>%
    select(Year, Mo, Globe) %>%
    mutate_all(as.numeric) %>% as_tibble() %>%
    rename(!!new_globe_name := Globe)
    )
  }

df_Lower_Troposphere <- clean_data(df_Lower_Troposphere_raw,
                                   new_globe_name = "Globe_Temp_Lower_Trop")
df_Mid_Troposphere <- clean_data(df_Mid_Troposphere_raw,
                                   new_globe_name = "Globe_Temp_Mid_Trop")
df_Tropopause <- clean_data(df_Tropopause_raw,
                                   new_globe_name = "Globe_Temp_Tropopause")
df_Lower_Stratosphere <- clean_data(df_Lower_Stratosphere_raw,
                                   new_globe_name = "Globe_Temp_Lower_Strato")

#combinding the 4 df to one
df_list <- list(df_Lower_Troposphere, 
            df_Mid_Troposphere[!colnames(df_Mid_Troposphere) %in% c("Mo")],
            df_Tropopause[!colnames(df_Tropopause) %in% c("Mo")], 
            df_Lower_Stratosphere[!colnames(df_Lower_Stratosphere) %in% c("Mo")])
  
df <- df_list %>% reduce(full_join,by='Year')
```

Calculate the 12-month(right-aligned) moving average

```{r}

for(col_name in colnames(df[,c(-1,-2)])){
  new_name = paste0(col_name,"_mov_av_right_align")
  df<- transform(df, placeholder_name = round(rollmean(df[col_name], k=3, fill = "extend",align = "right"),4))
  names(df)[names(df)==paste0(col_name,".1")]<-new_name
}
```

Calculate the average of the "moving average"

```{r}
df <- df %>%
  mutate(Avarage_moving_average = rowSums(df[7:10])/4)
```

Pivot to easier create graph

```{r}
df_pivot <- df %>% pivot_longer(!c(Year,Mo), names_to = "Type",values_to = "Temp")

df_pivot[df_pivot == "Globe_Temp_Lower_Trop_mov_av_right_align"] <- "Lower Troposphere"
df_pivot[df_pivot == "Globe_Temp_Mid_Trop_mov_av_right_align"] <- "Mid Troposphere"
df_pivot[df_pivot == "Globe_Temp_Tropopause_mov_av_right_align"] <- "Tropopause"
df_pivot[df_pivot == "Globe_Temp_Lower_Strato_mov_av_right_align"] <- "Lower Stratosphere"
df_pivot[df_pivot == "Avarage_moving_average"] <- "Avarage"
```

Draw Graph

```{r fig.asp = 0.6, fig.height=1, fig.width=8}
type_filter = c("Lower Troposphere",
                "Mid Troposphere",
                "Tropopause",
                "Lower Stratosphere",
                "Avarage")

plot_title = "Atmospheric Temperature"
plot_subtitle ="12th month moving average (right-aligned)"
plot_label_y = "Temprature"
plot_label_x = "Year"
plot_caption = "Source: https://www.drroyspencer.com/latest-global-temperatures (06/02/23)"
plot_leg_title = "Area"

suppressWarnings({
df_pivot %>% 
  filter(Type %in% type_filter) %>%
  ggplot(aes(x=Year,
             y=Temp,
             color = Type))+
  geom_line(size=0.8,linetype = "solid")+
  scale_color_manual(values=hod_theme_colours)+
  labs(title = plot_title,
       subtitle = plot_subtitle,
         y = plot_label_y,
         x = plot_label_x,
         caption = plot_caption)+
 scale_x_continuous(n.breaks = 9)+
  hod_theme(ledgend_pos="right", alter_x_axis_colour= TRUE)+
  theme(legend.margin=margin(t = -2,r=-0.2, unit='cm'),
        plot.caption = element_text(hjust=-120,vjust=-3))+
  guides(colour=guide_legend(title = plot_leg_title))
})
```
