---
title: "Task 2 sok1005"
format: html
editor: visual
---

## Task

1.  Use R code to download from [FRED](https://fred.stlouisfed.org/series/DCOILBRENTEU) the daily Brent Crude Oil Price from January 4th 2010 to December 30, 2022.

2.  Perform some data wrangling and convert the daily price to monthly price.

3.  Plot the monthly price against the date

**Setup**

```{r}
rm(list = ls())
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(fredr))


# A custome theme I use

hod_theme_colours <- c("#ffa822","#1ac0c6","#ff6150","#30B66A","#B06AFF","#FF21E1")

hod_theme <- function(alter_x_axis_colour = FALSE,
                      ledgend_pos = "right") {
  
  colour_text <- "#b8aba7"
  colour_text_light <- "#FFE9E3"
  colour_box <- "#2b2b2b"
  colour_lines <- "#827974"
  colour_main <- "#E9D5DA"
  colour_background <- "#2b2b2b"
  colour_palette_standard <- c("#ffa822", "#ff6150","#1ac0c6")
  colour_text_x_axis <- colour_text
  
  if(alter_x_axis_colour){
     colour_text_x_axis <- c(colour_text,colour_text_light)
  }
    
  theme(
    # Border and margin
    plot.margin = margin(0.4, 0.6, 0.4, 0.4, "cm"),
    panel.border = element_rect(colour = colour_box, fill = NA, linetype = 2),
    
    # background
    panel.background = element_rect(fill = colour_background),
    plot.background = element_rect(fill = colour_background),
    strip.background = element_rect(fill = colour_background),
    
    rect = element_rect(fill = colour_background, colour = "white",
                        size = 0.5, linetype = 1),
    
    # grid
    panel.grid.major.x = element_line(colour = colour_lines, linetype = 5, size = 0.25),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y =  element_line(colour = colour_lines, linetype = 5, size = 0.25),
    panel.grid.minor.y = element_blank(),
    axis.ticks = element_line(colour = colour_text),
    
    # line
    line = element_line(colour = colour_palette_standard),
    
    # text
    text = element_text(colour = colour_text, face = "italic", family = "mono"),
    plot.title = element_text(colour = colour_text_light, face = "bold", family = "mono"),
    plot.subtitle = element_text(colour = colour_text, family = "mono", hjust = 0.01),
    plot.caption = element_text(colour = colour_text, face = "italic", family = "mono"),
    axis.text.y = element_text(colour = colour_text, face = "italic", family = "mono"),
    axis.text.x = element_text(colour = colour_text_x_axis, face = "italic", family = "mono"),
    axis.title = element_text(colour = colour_text_light, family = "mono"),
    plot.tag = element_text(colour = colour_text, face = "italic", family = "mono"),
    legend.text = element_text(color = colour_text,face = "italic", family = "mono"),
    legend.title = element_text(colour = colour_text, face = "italic", family = "mono"),
    strip.text = element_text(colour = colour_text_light, face = "bold", family = "mono"),
    
    # legend
    legend.position = ledgend_pos,
    legend.key = element_rect(fill = colour_background)
  )
}
```

**Importing and cleaning of data**

### IMPORTANT: As the website FRED specified I removed my API key, and you have to put in your own to make the code work.

follow this link for more info: https://fred.stlouisfed.org/docs/api/api_key.html

```{r}
api_key = "" #IMPORTANT: API KEY FORM FRED
fredr_set_key(api_key)

if (fredr_has_key()) {
  df_raw <- fredr(
    series_id = "DCOILBRENTEU",
    observation_start = as.Date("2010-01-04"),
    observation_end = as.Date("2022-12-30"),
    frequency = "m",
    aggregation_method = "avg")
}
```

```{r}
df = df_raw %>%
  select(-c("realtime_start","realtime_end","series_id"))

df <- na.omit(df)
```

```{r}
df_month <- df %>% 
  mutate(date = as.Date(date, format = "%Y/%d/%m"))
```

**Plot**

```{r fig.asp = 0.6, fig.height=1, fig.width=8}

plot_title = "Crude Oil Prices"
plot_subtitle ="Average per month (2010-2022)"
plot_label_y = "Price"
plot_label_x = "Year"
plot_caption = "Source: https://fred.stlouisfed.org/series/DCOILBRENTEU (accessed: 19/02/23)"

suppressWarnings({
df_month %>% 
  ggplot(aes(x=date,
             y=value))+
  geom_line(size=0.8,linetype = "solid", color = hod_theme_colours[1])+
  labs(title = plot_title,
       subtitle = plot_subtitle,
         y = plot_label_y,
         x = plot_label_x,
         caption = plot_caption)+
  hod_theme()
})
```
